---
title: "project1"
author: "me"
date: "2021/7/12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Loading the packages
```{r}
library(dplyr)
library(ggplot2)
```
###Loading the data
```{r}
data <- read.csv(file.choose(), header=TRUE, sep=",")
```
###Subsetting the data
```{r}
data$date<-as.Date(data$date)
```
##What is mean total number of steps taken per day?
###Plot the histogram
```{r}
stepsPerDay <- data %>%
        group_by(date) %>%
        summarize(sumsteps = sum(steps, na.rm = TRUE)) 
hist(stepsPerDay$sumsteps, main = "Daily Steps", 
     col="green", xlab="Steps", ylim = c(0,30))
```
###Calculate the mean and median
```{r}
mean <- round(mean(stepsPerDay$sumsteps),digits = 2)
median <- round(median(stepsPerDay$sumsteps),digits = 2)
print(paste("The mean is", mean))
print(paste("The median is", median))
```
##What is the average daily activity pattern?
###Plot
```{r}
stepsPerInterval <- data %>%
        group_by(interval) %>%
        summarize(meansteps = mean(steps, na.rm = TRUE)) 
barplot(stepsPerInterval$meansteps ~ stepsPerInterval$interval,
     xlab = "Intervals", ylab = "Average Number of Steps",
     main = "Steps By Time Interval")
```
###Maximum number of step
```{r}
print(paste("Interval containing the most steps on average: ",stepsPerInterval$interval[which.max(stepsPerInterval$meansteps)]))
```
##Imput missing values
###Number of missing values
```{r}
print(paste("The total number of rows with NA is: ",sum(is.na(data$steps))))
```
###Replace NA
by replacing them with the mean
###Create new dataset
```{r}
newdata <- data
for (i in 1:nrow(data)){
        if(is.na(data$steps[i])){
                newdata$steps[i]<- stepsPerInterval$meansteps[newdata$interval[i] == stepsPerInterval$interval]
        }
}
```
###Plot the histogram+analysis
```{r}
stepsPerDay <- newdata %>%
        group_by(date) %>%
        summarize(sumsteps = sum(steps, na.rm = TRUE)) 
hist(stepsPerDay$sumsteps, main = "Histogram of Daily Steps", 
     col="green", xlab="Steps")
meanPostNA <- round(mean(stepsPerDay$sumsteps), digits = 2)
medianPostNA <- round(median(stepsPerDay$sumsteps), digits = 2)
print(paste("The mean is: ", mean(meanPostNA)))
print(paste("The median is: ", median(medianPostNA)))
```
According to the analysis, both mean and median becomes larger.
##Are there differences in activity patterns between weekdays and weekends?
###Create new variables
```{r}
activityDoW <- newdata
activityDoW$date <- as.Date(activityDoW$date)
activityDoW$day <- ifelse(weekdays(activityDoW$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
activityDoW$day <- as.factor(activityDoW$day)
```
###Plot
```{r}
activityWeekday <- filter(activityDoW, activityDoW$day == "weekday")
activityWeekend <- filter(activityDoW, activityDoW$day == "weekend")

activityWeekday <- activityWeekday %>%
        group_by(interval) %>%
        summarize(steps = mean(steps)) 
activityWeekday$day <- "weekday"

activityWeekend <- activityWeekend %>%
        group_by(interval) %>%
        summarize(steps = mean(steps)) 
activityWeekend$day <- "weekend"

wkdayWkend <- rbind(activityWeekday, activityWeekend)
wkdayWkend$day <- as.factor(wkdayWkend$day)


g <- ggplot (wkdayWkend, aes (interval, steps))
g + geom_line() + facet_grid (day~.) + 
        theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14)) + 
        labs(y = "Number of Steps") + labs(x = "Interval") + 
        ggtitle("Average Number of Steps - Weekday vs. Weekend") + 
        theme(plot.title = element_text(hjust = 0.5))
print(g)
```
